Global patterns of forest autotrophic carbon fluxes

### 2020-08-06 manuscript/manuscript draft.Rmd
working on discussion

**update this paragraph ([issue 78](https://github.com/forc-db/Global_Productivity/issues/78))** 
*Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$  $BNPP_{fine.root)$; Fig. 2; Tables S2, S6).* 
There are two, non-exclusive, potential explanations for this. 
First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes. 
Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may use fundamentally different approaches in different sites (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{fine.root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{root}$), and sampling depth is variable and often insufficient to capture the full soil profile. 
$ANPP_{stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled [@clark_measuring_2001]. 
However, methodological variation and uncertainty affect all of fluxes considered here, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes. 
Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. 
That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors (**REFS**).

*At least at this global scale, we find little evidence of any non-linearity in temperature's influence on C flux.* 
The latitudinal gradient in forest C flux (Fig. 2) is driven primarily by temperature-related climate variables, the effects of which are moderated by moisture availability (Table 1, *H2-H3*; Figs. 3-4).  MAT and MAP have long been identified as primary global-scale drivers of C fluxes [@lieth_primary_1973; **REFS**; @taylor_temperature_2017]. It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of C fluxes because many climate variables co-vary across the latitudinal gradient (Fig. S2), because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means. However, it remains informative to consider these relationships. Among the temperature-related climate variables, $MAT$ is generally the most strongly correlated with C fluxes (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products, introducing spatial scaling errors.

This finding supports the continued focus on MAT as a primary--*albiet* not mechanistic--correlate of C fluxes. The effects of $MAT$ are modified by moisture availability, with reduced C fluxes under relatively dry conditions (*i.e.*, low precipitation; high vapour pressure deficit) and sometimes under very high precipitation (Figs. 3-4). **BBL: I feel like this is pretty repetitive of previous paragraph.** The observed positive interaction between MAT and MAP for $ANPP_{stem}$ on the global scale (Fig. 3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20$\circ$C [@taylor_temperature_2017]. *(@sullivan_sensitivity_2020 show a decline in $ANPP_{stem}$ in sites with high max temperatures and limited moisture.)* However, we detect no such interaction for $ANPP$ or most other C fluxes, and we find a contrasting negative interaction for $NPP$ (Fig. 3), suggesting that more data are required to sort out potential differences in the interactive effects of MAT and MAP on C fluxes in the tropics.

Forest autotrophic C fluxes decline with temperature seasonality (Table 1, *H4*; Fig. 4e), and are minimal during cold- or dry- dormant seasons. To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; @anderson_temperature-dependence_2006; @enquist_adaptive_2007; @michaletz_convergence_2014]. The sort of simple metric that has been used to define growing season at a global scale [@kerkhoff_plant_2005] is coarse with respect to temperature because it's calculated on a monthly timescale and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow. We found that a temperature-defined growing season length had strong positive correlation with C fluxes (Fig. 4f), but was never the best. Dividing annual fluxes by growing season length to yield average flux per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007]. However, there remained a number of significant correlations with growing season climatic conditions, indicating that climatic conditions remain influential within the growing season. We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power. Mechanistic accounting for climatic effects on global forest carbon flux patterns instead requires models representing physiologically meaningful timescales [e.g., **REFS**; @longo_biophysics_2019].

Our analysis clarifies how forest autotrophic carbon fluxes vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming. 
Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of forest C cycling (Figs. 2-3).
This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018]. 
However, increasing C flux rates are by no means universal [e.g., @rutishauser_testing_2020; @hubau_asynchronous_2020], likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.  
In addition, some tropical forests are
**(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)** [@sullivan_sensitivity_2020]
Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate. 
Nevertheless, understanding of the fundamental climatic controls on C cycling sets a firmer foundation for understanding forest C cycle responses to accellerating climate change.

>>>>

*Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$  $BNPP_{fine.root)$; Fig. 2; Tables S2, S6).* 
There are two, non-exclusive, potential explanations for this. 
First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes. 
Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may use fundamentally different approaches in different sites (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{fine.root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{root}$), and sampling depth is variable and often insufficient to capture the full soil profile. 
$ANPP_{stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled [@clark_measuring_2001]. 
However, methodological variation and uncertainty affect all of fluxes considered here, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes. 
Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. 
That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors (**REFS**).

*Temperature and its seasonality were the primary drivers of C flux on the global scale (Figs. 2-4), consistent with a long legacy of research identifying temperature as a primary driver of forest ecosystem C cycling [e.g., @lieth_primary_1973; **REFS].*
We find little evidence of any non-linearity in temperature's influence on C flux.
The relationship of all fluxes to $MAT$ as an individual driver were best described by a linear function (Table S2) - with the exception of $BNPP$, whose response to $MAT$ was close to linear (Fig. 4a).
This result contrasts with both the idea that C flux responds exponentially to temperature (**REFS**) and with the idea that fluxes saturate with $MAT$ below approximately 25$^\circ$C MAT (**REFS**).
It remains possible that flux declines above this threshold [@larjavaara_temperature_2012; @sullivan_sensitivity_2020], as is also consistent with tree-ring records indicating that tropical tree growth declines at high temperatures [e.g., @vlam_temperature_2014].
However, these higher temperatures also tend to be associated with drier conditions (*i.e.*, low precipitation, high PET and VPD), all of which are associated with reduced C fluxes (Figs. 4c-d, S4-S5).
Indeed, while temperature responses dominate at this global scale and with the climate space occupied by forests, the effects of temperature are moderated by moisture availability (Table 1; Figs 3-4).
MAT and MAP have long been identified as primary global-scale drivers of C fluxes [@lieth_primary_1973; **REFS**; @taylor_temperature_2017].

*It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of C fluxes because many climate variables co-vary across the latitudinal gradient (Fig. S2), because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means. However, it remains informative to consider these relationships. Among the temperature-related climate variables, $MAT$ is generally the most strongly correlated with C fluxes (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products, introducing spatial scaling errors.*

This finding supports the continued focus on MAT as a primary--*albiet* not mechanistic--correlate of C fluxes. The effects of $MAT$ are modified by moisture availability, with reduced C fluxes under relatively dry conditions (*i.e.*, low precipitation; high vapour pressure deficit) and sometimes under very high precipitation (Figs. 3-4). **BBL: I feel like this is pretty repetitive of previous paragraph.** The observed positive interaction between MAT and MAP for $ANPP_{stem}$ on the global scale (Fig. 3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20$\circ$C [@taylor_temperature_2017].  However, we detect no such interaction for $ANPP$ or most other C fluxes, and we find a contrasting negative interaction for $NPP$ (Fig. 3), suggesting that more data are required to sort out potential differences in the interactive effects of MAT and MAP on C fluxes in the tropics.

Forest autotrophic C fluxes decline with temperature seasonality (Table 1, *H4*; Fig. 4e), and are minimal during cold- or dry- dormant seasons. 
To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; @anderson_temperature-dependence_2006; @enquist_adaptive_2007; @michaletz_convergence_2014]. 
The sort of simple metric that has been used to define growing season at a global scale [@kerkhoff_plant_2005] is coarse with respect to temperature because it is calculated on a monthly timescale and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow. 
We found that a temperature-defined growing season length had strong positive correlation with C fluxes (Fig. 4f), but was never the best univariate predictor. 
Dividing annual fluxes by growing season length to yield average flux per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007]. 
However, there remained a number of significant correlations with growing season climatic conditions, indicating that climatic conditions remain influential within the growing season (Figs. S8-S9). 
We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power. 
Mechanistic accounting for climatic effects on global forest carbon flux patterns instead requires models representing physiologically meaningful timescales [e.g., @medvigy_mechanistic_2009; @longo_biophysics_2019].

Our analysis clarifies how forest autotrophic carbon fluxes vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming. 
To the extend that patterns across broad scale climatic gradients can foretell how ecosystem responses to climate change, our findings suggest that higher temperatures with similar moisture availability result in a generalized acceleration of forest C cycling (Figs. 2-3).
This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018]. 
However, increasing C flux rates are by no means universal [e.g., @rutishauser_testing_2020; @hubau_asynchronous_2020], likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.  
Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate. 
Indeed, tree-ring studies from forests around the world indicate that tree growth rates - along with $ANPP_{stem}$ and possibly other ecosystem C fluxes - respond negatively to temperature [@sniderhan_growth_2016;@helcoski_growing_2019].
Furthermore, in the tropics, climate change will push forests beyond any contemporary climate, and there are some indications that this could reduce C flux rates  [@sullivan_sensitivity_2020] (**REFS**).
Thus, while further research is required to understand the extent to which forest responses to climate change will track global gradients, and the time scale on which they will do so, understanding the fundamental climatic controls on annual C cycling in Earth's forests sets a firmer foundation for understanding forest C cycle responses to accelerating climate change.
